---
title: "Assignment - 1"
author: "Akhil Gaur"
output: html_document
---

-----------------------------------------
# Reading the ODI Batting Analysis File 


```{r}
odi <- read.csv("C:/Users/Admin/Desktop/odi-batting-analysis.csv", sep = ",")
head(odi)
```

# Bar chart for total runs by month 


```{r}
library(dplyr)
library(ggplot2)
month_runs <- odi %>% group_by(Month) %>% summarise(TotalRuns = sum(Runs, na.rm = TRUE))
month_runs$Month <- factor(month_runs$Month)
ggplot(month_runs, aes(x = Month, y = TotalRuns,fill = -TotalRuns)) + geom_bar(stat = 'identity',color = "pink") +coord_flip() 
```

# Bar chart for total runs by day of week 


```{r}
weekday_runs <- odi %>% group_by(Dayofweek) %>% summarise(TotalRuns = sum(Runs, na.rm = TRUE))
weekday_runs$Dayofweek <- factor(weekday_runs$Dayofweek)
ggplot(weekday_runs, aes(x = Dayofweek, y = TotalRuns,fill = -TotalRuns)) + geom_bar(stat = 'identity',color = "brown")+coord_flip()
```


# Bar chart for total runs by day of month 


```{r}
day_month_run <- odi %>% group_by(Day) %>% summarise(TotalRuns = sum(Runs, na.rm = TRUE))
day_month_run$Day = factor(day_month_run$Day)
ggplot(day_month_run, aes(x = Day, y = TotalRuns,fill = -TotalRuns)) + geom_bar(stat = 'identity')+coord_flip()
```

----------------------------------------------------------------------------------------
#2. Create a new column by binning the Runs column using the following ranges and plot the  frequency count of ranges
# a. 0-25,
# b. 26-50,
# c. 51-75,
# d. 76-100,
# e. 100+ 
-----------------------------------------------------------------------------------------
```{r}
odi <- na.omit(odi)
odi$bin <- cut(odi$Runs, breaks = c(0,25,50,75,100,200), labels = c("0-25", "26-50", "51-75", "76-100", "100+"), include.lowest = TRUE)
ggplot(odi, aes(x = bin)) + geom_bar(fill = "green")+coord_flip()
```

----------------------------------------------------------------------------------------
#3. Top ten players by total runs 
----------------------------------------------------------------------------------------

```{r}
players_summary <- odi %>% group_by(Player) %>% summarise(TotalRuns = sum(Runs, na.rm = TRUE))
top10_players_summary <- head(players_summary %>% arrange(-TotalRuns), 10)
top10_players_summary
ggplot(top10_players_summary, aes(x = Player, y = TotalRuns, fill = -TotalRuns)) + geom_bar(stat = "identity") + coord_flip()
```


# Number of centuries across years 


```{r}
a<-odi %>% filter(odi$Player %in% top10_players_summary$Player,Runs >= 100)
centuries_years <- a %>% group_by(Player,Year) %>% summarise(Centuries = n())
ggplot(centuries_years, aes(x = Year, y = Centuries)) + geom_bar(stat = "identity", fill = "blue")+facet_wrap(~Player)

```

# Number of ducks across years 


```{r}
b<- odi %>% filter(odi$Player %in% top10_players_summary$Player & Runs == 0)
ducks_years <- b %>% group_by(Player,Year) %>% summarise(Ducks = n())
ggplot(ducks_years, aes(x = Year, y = Ducks)) + geom_bar(stat = "identity",fill="red")+ facet_wrap(~Player)
```

----------------------------------------------------------------------------------------
#4a. Calculating total number of runs by each player and then filter only for those players who have scored more than 5000 runs.
----------------------------------------------------------------------------------------

```{r}
players_summary_total <- odi %>% group_by(Player) %>% summarise(TotalRuns = sum(Runs)) %>% arrange(TotalRuns)
runs_5000 <- filter(players_summary_total, TotalRuns >= 5000)
h<-head(runs_5000,10)
ggplot(h,aes(x=Player,y=TotalRuns,fill = -TotalRuns)) + geom_bar(stat='identity')+coord_flip()

```

-----------------------------------------------------------------
#4b. Filter only those players who have played more than 50 matches.
-----------------------------------------------------------------

```{r}
players_summary_matches <- odi %>% group_by(Player) %>% summarise(TotalRuns = sum(Runs),count = n()) %>% arrange(TotalRuns)
matches_50 <- filter(players_summary_matches, count > 50)
h1<-head(matches_50,10)
ggplot(h1,aes(x=Player,y=TotalRuns,fill = -TotalRuns)) + geom_bar(stat='identity')+coord_flip()

```

--------------------------------------------------------------------------------
#5. Create a new column by computing the number of words in each palyers's name.
--------------------------------------------------------------------------------

```{r}
odi$Player<-as.character(odi$Player)

c<-strsplit(odi$Player," ")
odi$Freqwords<-sapply(c,length)
odi$Freqwords
str(odi)

```

---------------------------------------------------------------------------------
#6. Crosstab between month and day.
---------------------------------------------------------------------------------

```{r}
x<-xtabs(~Month+Day,data=odi)
ftable(x)

```